---
title: "Installation instructions for hicream"
author: "Elise Jorge, Sylvain Foissac, Pierre Neuvial, and Nathalie Vialaneix"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Installation instructions for hicream}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# AWK

`hicream` requires the `awk` command to be installed on your system. If you are 
using a Unix-based system, `awk` is likely already installed. If you are using 
Windows a [cywgin](https://www.cygwin.com/) version of `awk` is searched for by 
the package. You can install `awk` by downloading the Gawk package from the GNU
website: https://www.gnu.org/software/gawk/.

More about AWK: https://en.wikipedia.org/wiki/AWK


# Python dependencies

`hicream` requires the `python` command to be installed on your system (>=3.9).
Additional Python packages (`scikit-learn`, `kneebow`, `numpy`, and `pandas`) 
are also required.

More about used Python packages:

* scikit-learn: https://scikit-learn.org/stable/
* kneebow: https://pypi.org/project/kneebow/
* numpy: https://numpy.org/
* pandas: https://pandas.pydata.org/

To install Python modules, please check "Installing Python modules" in 
[Python documentation](https://docs.python.org). With pip:
```
pip install scikit-learn
pip install kneebow
pip install numpy
pip install pandas
```

# Session information

```{r}
sessionInfo()
```

